---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# dyRank

<!-- badges: start -->
<!-- badges: end -->

The goal of `dyRank` is to estimate rating of players based on rank-ordered data 
with time-index.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("sou412/dyRank")
```


## Example: Single Ranking Type


As an example, we use `f1_race` data in the `dyRank` package. 

```{r echo=TRUE, message=FALSE, warning=FALSE}
## load package 
require(dyRank)
require(tidyverse)
```


```{R}
## load data 
data("f1_race", package = "dyRank")

## view data 
f1_race
```


### Estimate rating 


```{r eval=FALSE}
## estimate via MCMC 
fit <- dyRank(
	data       = f1_race, 
	var_rank   = "Pos",           
	var_player = "driver", 
	var_match  = "GP",
	var_time   = "year",
	driver_fix = "Aguri Suzuki",
	mcmc = 100, burnin = 10, thin = 1,
	truncation = 3
)
```


**Arguments**:

- `data`: A data frame that contains variables specified in `var_rank`, `var_player`, `var_match`, `var_time`.
		This should be an object of either `data.frame` class or `tibble` class.
+ `var_rank`: A variable name (in character) of the outcome that records the ranking. The top rank should take `1`, and all ranking values within a match (specified in `var_match`) should be consecutive. `NA` values is allowed. 
+ `var_player`: A variable name (in character) of the players. 
+ `var_match`: A variable name (in character) of the matches. The package allows for the panel strucutre, which menas that the same match can repeat over time. 
+ `var_time`: A variable name of time index. 
+ `driver_fix`: A name (or index) within `var_player` that is used as a refenrece. This should be chosen carefully, otherwise estimates will be unstable. Generally, a player who appears in multiple matches and years and who can be considered as a "middle player" is recommended as a reference. 
+ `mcmc`, `burnin` and `thin` correspond to the MCMC parameters that specify the iterations, burin periods and the thinning, respectively. Default values are set small; in practice longer values are recommended.
+ `truncation`: A truncation parameter that takes `1` or larger. This parameter specifies if we want to ignore players who ranked `k` or lower where `k` is the vlaue of `truncation`. The value of `1` corresponds to the original Placket-Luce representation, but the estimation might not be stable when observations are dropped due to `NA` values. 



