---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# dyRank: Dynamic Rating Estimation for Ranked Data

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/soichiroy/dyRank.svg?branch=master)](https://travis-ci.org/soichiroy/dyRank)
<!-- badges: end -->

The goal of `dyRank` is to estimate the dynamic rating of items, players or students based on rank-ordered data 
with time-index.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("soichiroy/dyRank")
```


## Example: Single Ranking Type


As an example, we use `f1_race` data in the `dyRank` package. 

```{r echo=TRUE, message=FALSE, warning=FALSE}
## load package 
require(dyRank)
require(tidyverse)
```


```{R}
## load data 
data("f1_race", package = "dyRank")

## see ?f1_race for the details of this dataset.

## view data 
f1_race
```


### Estimate rating 


```{r eval=TRUE, cache=TRUE}
## set seed
set.seed(1234)

## estimate via MCMC 
fit <- dyRank(
	data       = f1_race, 
	var_rank   = "Pos",           
	var_player = "driver", 
	var_match  = "GP",
	var_time   = "year",
	driver_fix = "Timo Glock",
	mcmc = 100, burnin = 10, thin = 1,
	truncation = 3
)
```


**Arguments**:

- `data`: A data frame that contains variables specified in `var_rank`, `var_player`, `var_match`, `var_time`.
		This should be an object of either `data.frame` class or `tibble` class.
+ `var_rank`: A variable name (in character) of the outcome that records the ranking. The top rank should take `1`, and all ranking values within a match (specified in `var_match`) should be consecutive. `NA` values is allowed. 
+ `var_player`: A variable name (in character) of the players. 
+ `var_match`: A variable name (in character) of the matches. The package allows for the panel strucutre, which menas that the same match can repeat over time. 
+ `var_time`: A variable name of time index. 
+ `driver_fix`: A name (or index) within `var_player` that is used as a refenrece. This should be chosen carefully, otherwise estimates will be unstable. Generally, a player who appears in multiple matches and years and who can be considered as a "middle player" is recommended as a reference. 
+ `mcmc`, `burnin` and `thin` correspond to the MCMC parameters that specify the iterations, burin periods and the thinning, respectively. Default values are set small; in practice longer values are recommended.
+ `truncation`: A truncation parameter that takes `1` or larger. This parameter specifies if we want to ignore players who ranked `k` or lower where `k` is the vlaue of `truncation`. The value of `1` corresponds to the original Placket-Luce representation, but the estimation might not be stable when observations are dropped due to `NA` values. 

**Return**

`dyRank()` returns a list of two elements.

- `lambda`: Estimated parameters stored as a list. Each element of the list corresponds to an interation 
	of the MCMC step. 
	
```{r}
## total number of elements
length(fit$lambda)

## each element of lambda is a list of estimates for all drivers
length(fit$lambda[[1]])

## each element of lambda[[k]] is an estimate for drivers
length(fit$lambda[[1]][[1]])
```
- `data` returns the list of formatted datasets.

```{r}
## formatted data used for the estimation 
fit$data$dat_ref

## global information
fit$data$n_drivers
fit$data$n_race
```

### Obtaining the summary estimates 


```{r}
## get the summary of rating 
rating <- get_rating(fit)

## view the estimates 
rating
```

```{r, fig.width=7, fig.height=3}
## example visualization 
drivers_use <- c("Michael Schumacher", "Lewis Hamilton", "Sebastian Vettel", "Kimi Räikkönen")

## plot rating with plot_rating()
gg <- plot_rating(rating, facet = TRUE, ncol = 4, driver_name = drivers_use)
gg + ylim(-2, 8) + xlim(1984, 2019)
```


### Assessing estimates and convergence


```{r}
## load coda package 
require(coda)

## convert estimates to the MCMC object 
mcmc_obj <- get_mcmc(fit)

## who is the first driver in the data?
names(mcmc_obj)[1]

## plot the first driver's rating estimate 
plot(mcmc_obj[[1]])
```


#### Checking covergence with Geweke statistics 

```{r fig.height=4, fig.width=7}
## plot the autocorrelation 
autocorr.plot(mcmc_obj[[1]])

## geweke plot 
geweke.plot(mcmc_obj[[1]])
```


### Running with Multiple Chains 


```{r cache=TRUE, message=FALSE, warning=FALSE}
## load additional package for parallel
require(furrr)

## setup parallel 
plan(multiprocess)

## run with multiple chains 
n_chains <- 3
set.seed(1234)
fit_nchains <- future_map(1:n_chains, function(chains) {
	fit_tmp <- dyRank(
		data       = f1_race, 
		var_rank   = "Pos",           
		var_player = "driver", 
		var_match  = "GP",
		var_time   = "year",
		driver_fix = "Timo Glock",
		mcmc = 100, burnin = 10, thin = 1,
		truncation = 3
	)
}, .options = future_options(seed = TRUE))
```

#### Gelman-Rubin statistic

```{r, fig.height=4, fig.width=7}
## summarize results 
d1_mcmc <- map(fit_nchains, ~get_mcmc(.x))

## gelman rubin statistics 
coda::gelman.plot(
	mcmc.list(d1_mcmc[[1]][[1]], d1_mcmc[[2]][[1]], d1_mcmc[[3]][[1]])
)
```

## Example: Multiple Ranking Types


```{r}
## load additional data 
## see ?f1_grid and ?f1_laptime for the details 
data("f1_grid", package = "dyRank")
data("f1_laptime", package = "dyRank")

## prepare data 
f1_all <- bind_rows(f1_race, f1_grid, f1_laptime)
```



### Estimate rating via `hdyRank()`



```{r cache=TRUE}
## fit hierarhcal model
set.seed(1234)
fit_hier <- hdyRank(
	data       = f1_all,
	var_rank   = "Pos",           
	var_player = "driver", 
	var_match  = "GP",
	var_time   = "year",
	var_rank_type = "rank_type",
	driver_fix = "Timo Glock",
	mcmc = 100, burnin = 10, thin = 1, 
	truncation = 5
)
```


**Arguments**

- `hdyRank()` inherits all the arguments specified in `dyRank()`.
- It takes the additional argument `var_rank_type`: A variable name (in character) of the rank type. When this variable takes a single level (i.e., only one type of ranking), `dyRank()` should be used instead.



### Estimated rating 


```{r, fig.width=7, fig.height=3}
# obtain rating 
rating_hier <- get_rating(fit_hier)

## example visualization 
drivers_use <- c("Michael Schumacher", "Lewis Hamilton", "Sebastian Vettel", "Kimi Räikkönen")

## plot rating with plot_rating()
gg <- plot_rating(rating_hier, facet = TRUE, ncol = 4, driver_name = drivers_use)
gg + ylim(-2, 8) + xlim(1984, 2019)
```