---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# dyRank: Dynamic Rating Estimation for Ranked Data

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/soichiroy/dyRank.svg?branch=master)](https://travis-ci.org/soichiroy/dyRank)
[![R build status](https://github.com/soichiroy/dyRank/workflows/R-CMD-check/badge.svg)](https://github.com/soichiroy/dyRank/actions)
<!-- badges: end -->

The goal of `dyRank` is to estimate the dynamic rating of items, players or students based on rank-ordered data
with time-index.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("soichiroy/dyRank")
```


## Example: Single Ranking Type


As an example, we use `f1_race` data in the `dyRank` package.

```{r echo=TRUE, message=FALSE, warning=FALSE}
## load package
require(dyRank)
require(tidyverse)
require(knitr)
```


```{R}
## load data
## see ?f1_race for the details of this dataset
data("f1_race", package = "dyRank")
```


### Estimate rating via `dyRank()`


```{r eval=TRUE, cache=TRUE}
## set seed
set.seed(1234)

## estimate via MCMC
fit <- dyRank(
	data       = f1_race,
	var_rank   = "Pos",           
	var_player = "driver",
	var_match  = "GP",
	var_time   = "year",
	driver_fix = "Timo Glock",
	mcmc = 100, burnin = 10, thin = 1,
	truncation = 3
)
```


**Arguments**:

| Argument | Description |
| -------- | ----------- |
| `data`   | A data frame that contains variables specified in `var_rank`, `var_player`, `var_match`, `var_time`. This should be an object of either `data.frame` class or `tibble` class. |
| `var_rank`| A variable name (in character) of the outcome that records the ranking. The top rank should take `1`, and all ranking values within a match (specified in `var_match`) should be consecutive. `NA` values is allowed. |
| `var_player`| A variable name (in character) of the players. |
| `var_match`| A variable name (in character) of the matches. The package allows for the panel structure, which means that the same match can repeat over time. |
| `var_time`| A variable name of time index. |
| `driver_fix`| A name (or index) within `var_player` that is used as a reference. This should be chosen carefully, otherwise estimates will be unstable. Generally, a player who appears in multiple matches and years and who can be considered as a "middle player" is recommended as a reference. |
| `mcmc`, | Number of MCMC iterations. | 
| `burnin`| Burn-in periods. `mcmc + burnin` iterations will run.| 
|`thin`   | Thinning periods. |
| `truncation`| A truncation parameter that takes `1` or larger. This parameter specifies if we want to ignore players who ranked `k` or lower where `k` is the value of `truncation`. The value of `1` corresponds to the original Placket-Luce representation, but the estimation might not be stable when observations are dropped due to `NA` values. |

**Return**

`dyRank()` returns a list of two elements.

- `lambda`: Estimated parameters stored as a list. Each element of the list corresponds to an iteration
	of the MCMC step.

```{r}
## total number of elements
length(fit$lambda)

## each element of lambda is a list of estimates for all drivers
length(fit$lambda[[1]])

## each element of lambda[[k]] is an estimate for drivers
length(fit$lambda[[1]][[1]])
```
- `data` returns the list of formatted datasets.

```{r}
## formatted data used for the estimation
dim(fit$data$dat_ref)

slice_head(as_tibble(fit$data$dat_ref), n = 5) %>%
	knitr::kable(digits = 3)


## global information
fit$data$n_drivers
fit$data$n_race
```

### Obtaining the estimated rating via `get_rating()`


```{r}
## get the summary of rating
rating <- get_rating(fit)

## view the estimates
slice_head(as_tibble(rating), n = 5) %>%
	knitr::kable(digits = 3)
```

```{r, fig.width=7, fig.height=3}
## example visualization
drivers_use <- c("Michael Schumacher", "Lewis Hamilton", "Sebastian Vettel", "Kimi Räikkönen")

## plot rating with plot_rating()
gg <- plot_rating(rating, facet = TRUE, ncol = 4, driver_name = drivers_use)
gg + ylim(-2, 8) + xlim(1984, 2019)
```


### Assessing convergence via `get_mcmc()` and `coda` package


```{r}
## load coda package
require(coda)

## convert estimates to the MCMC object
mcmc_obj <- get_mcmc(fit)

## who is the first driver in the data?
names(mcmc_obj)[1]

## plot the first driver's rating estimate
plot(mcmc_obj[[1]])
```


#### Checking covergence with Geweke statistics

```{r fig.height=4, fig.width=7}
## plot the autocorrelation
coda::autocorr.plot(mcmc_obj[[1]])

## geweke plot
coda::geweke.plot(mcmc_obj[[1]])
```


### Running with Multiple Chains


`dyRank` package provides a function `bind_chains()` to work with multiple chains.

- Store multiple chains as a list
- `bind_chains()` with `summarize = FALSE` behaves like `get_mcmc()`
- `bind_chains()` with `summarize = TRUE` behaves like `get_rating()`


```{r cache=TRUE, message=FALSE, warning=FALSE}
## load additional package for parallel
require(furrr)

## setup parallel
plan(multiprocess)

## run with multiple chains and store output as a list
n_chains <- 3
set.seed(1234)
fit_nchains <- future_map(1:n_chains, function(chains) {
	fit_tmp <- dyRank(
		data       = f1_race,
		var_rank   = "Pos",           
		var_player = "driver",
		var_match  = "GP",
		var_time   = "year",
		driver_fix = "Timo Glock",
		mcmc = 100, burnin = 10, thin = 1,
		truncation = 3
	)
}, .options = future_options(seed = TRUE))
```

#### Gelman-Rubin statistic

Multiple chains are requires to comute the Gelman-Rubin statistics (`gelman.diag()` and `gelman.plot()` in `coda` package).
`bind_chains()` returns a list of `mcmc.list` object where each element of the returned list corresponds to a `mcmc.list` object for each player (i.e., each level of `var_player`).



```{r, fig.height=4, fig.width=7}
## combine estimates into a single mcmc.list
m_list <- bind_chains(fit_nchains)

## gelman rubin statistics (for the first driver)
coda::gelman.plot(m_list[[1]])
```

#### Estimated rating (all chains combined)


We can obtain the summary statistics of the estimated rating by specifying `summarize = TRUE` option in `bind_chains()` function.
As `bind_chains()` with `summarize = TRUE` option returns an object of `dyRank.summary` class,
a user can use `plot_raing()` function to visualize the estimates as in the case of a single chain.


```{r}
## summarise estimates by combining chains
m_summary <- bind_chains(fit_nchains, summarize = TRUE)

## class
class(m_summary)
```

## Example: Multiple Ranking Types


```{r}
## load additional data
## see ?f1_grid and ?f1_laptime for the details
data("f1_grid", package = "dyRank")
data("f1_laptime", package = "dyRank")

## prepare data
f1_all <- bind_rows(f1_race, f1_grid, f1_laptime)
```



### Estimate rating via `hdyRank()`



```{r cache=TRUE}
## fit hierarhcal model
set.seed(1234)
fit_hier <- hdyRank(
	data       = f1_all,
	var_rank   = "Pos",           
	var_player = "driver",
	var_match  = "GP",
	var_time   = "year",
	var_rank_type = "rank_type",
	driver_fix = "Timo Glock",
	mcmc = 100, burnin = 10, thin = 1,
	truncation = 5
)
```


**Arguments**

- `hdyRank()` inherits all the arguments specified in `dyRank()`.
- It takes the additional argument `var_rank_type`: A variable name (in character) of the rank type. When this variable takes a single level (i.e., only one type of ranking), `dyRank()` should be used instead.



### Estimated rating


```{r, fig.width=7, fig.height=3}
# obtain rating
rating_hier <- get_rating(fit_hier)

## example visualization
drivers_use <- c("Michael Schumacher", "Lewis Hamilton", "Sebastian Vettel", "Kimi Räikkönen")

## plot rating with plot_rating()
gg <- plot_rating(rating_hier, facet = TRUE, ncol = 4, driver_name = drivers_use)
gg + ylim(-2, 8) + xlim(1984, 2019)
```
